The Birth of a Galaxy: Self-consistent transition to Population II Star Formation

### 2010-09-11 ms.tex
Changed subsection title.
Included all cited references.  Put bullet points for result subsections.

\shorttitle{POPULATION III STAR EPOCH}
\shortauthors{WISE ET AL.}

\title{How Short Lived is the Population III Epoch?}

@@

The transition from Pop III to Pop II star formation is solely
dependent on the propagation of metals from the SNe remnants into
future sites of star formation.  Their flows are complex because of
the interactions between the SN blastwave, cosmological accretion and
halo mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h}
\sim 10^6~\Ms$), radiation from the massive Pop III drives a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \citep{Kitayama04, Whalen04, Abel07}.  The latter
feature aids in the expansion of the blastwave because it delays the
transition to the Sedov-Taylor and snowplow phases.  In PISNe,
approximately half of the metals stay in the IGM with an metal bubble
size of a few kpc.  The remaining metals fall back into the host halo.
The blastwave may induce star formation in nearby halos through the
compression of the gas \citep{Ferrara98}; however timescales for metal
mixing into the dense gas are many dynamical times \citep{Cen07} for
shock velocities $\lsim 100~\kms$.

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman97, OShea2004} that employs our implementation
of adaptive ray tracing for radiation transport \citep{Wise10}.  To
resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior.  We also refine on a
DM overdensity of $3 \times 2^l$ and always resolve the local Jeans
length by at least 4 cells.  The latter criterion avoids any artifical
fragmentation during gaseous collapses \citep{Truelove97}.  We
initialize the simulation with \textsl{grafic} \citep{Bertschinger01}
at $z = 130$ and use the cosmological parameters from the 7-year WMAP
data \citep{WMAP7}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$,
$\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$
with the variables having their usual definitions.

@@

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{2009Sci...325..601T}
has recently shown that Pop III binaries may form in a fraction of
such halos.  Furthermore, detailed one-dimensional calculations have
shown that their IMF should follow a Kroupa-like IMF that has a
Salpeter slope at the high-mass end and an exponential cut-off below
$M_{\rm char}$ \citep{refs}.

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  The critical overdensity in which a star
forms is $5 \times 10^5$.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{ref} has been
called for use when restarting the simulations.

@@

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \citet{Schaerer02}.  The radiation field is evolved with
with adaptive ray tracing \citep{Abel02, Wise10} and is coupled
self-consistently to the hydrodynamics.  These stars die as
pair-instability SNe (PISNe) if they are in the mass range 140--260
\Ms \citep{Heger02}.  We use the explosion energy from
\citeauthor{Heger02}, where we fit the following function to their
models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He} -
64)]$, where $M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the
helium core mass and $M_\star$ is the stellar mass.

@@

\subsection{Metal enrichment history}

@@

@@

@@

J.H.W. is supported by the Hubble Fellowship etc.  The majority of the
analysis and plots were done with \texttt{yt} \citep{yt}.

>>>>

\shorttitle{THE BIRTH OF A GALAXY}
\shortauthors{WISE ET AL.}

\title{The Birth of a Galaxy: Self-consistent transition to Population
  II Star Formation}

@@

The transition from Pop III to Pop II star formation is solely
dependent on the propagation of metals from the SNe remnants into
future sites of star formation.  Their flows are complex because of
the interactions between the SN blastwave, cosmological accretion and
halo mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h}
\sim 10^6~\Ms$), radiation from the massive Pop III drives a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \citep{Kitayama04, Whalen04, Abel07}.  The latter
feature aids in the expansion of the blastwave because it delays the
transition to the Sedov-Taylor and snowplow phases.  In PISNe,
approximately half of the metals stay in the IGM with an metal bubble
size of a few kpc.  The remaining metals fall back into the host halo.
The blastwave may induce star formation in nearby halos through the
compression of the gas \citep{Ferrara98}; however timescales for metal
mixing into the dense gas are many dynamical times \citep{Cen08} for
shock velocities $\lsim 100~\kms$.

@@

We use the adaptive mesh refinement (AMR) code
\enzo~\citep{BryanNorman1997, OShea2004} that employs our
implementation of adaptive ray tracing for radiation transport
\citep{Wise10}.  To resolve minihalos with at least 100 dark matter
(DM) particles and follow the formation of the first generation of
dwarf galaxies, we use a simulation box of 1 Mpc that has a resolution
of $256^3$.  This gives us a DM mass resolution of 1840 \Ms.  We
refine the grid on baryon overdensities of $3 \times 2^{-0.2l}$, where
$l$ is the AMR level, resulting in a super-Lagrangian behavior.  We
also refine on a DM overdensity of $3 \times 2^l$ and always resolve
the local Jeans length by at least 4 cells.  The latter criterion
avoids any artifical fragmentation during gaseous collapses
\citep{Truelove97}.  We initialize the simulation with \textsl{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP data \citep{WMAP7}: $\Omega_M =
0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h = 0.71$,
$\sigma_8 = 0.81$, and $n = 0.963$ with the variables having their
usual definitions.

@@

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{2009Sci...325..601T}
has recently shown that Pop III binaries may form in a fraction of
such halos.

% Furthermore, detailed one-dimensional calculations have shown that
% their IMF should follow a Kroupa-like IMF that has a Salpeter slope
% at the high-mass end and an exponential cut-off below $M_{\rm char}$
% \citep{refs}.

For Pop III stars, we use the same star formation model as
\citet{Abel07} and \citet{Wise08_Gal} where each star particle
represents a single star.  The critical overdensity in which a star
forms is $5 \times 10^5$.  Instead of using a fixed stellar mass, we
randomly sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we initialize both
simulations with the same random seed and record the number of times
the random number generator \citep[Mersenne twister;][]{MTwister} has
been called for use when restarting the simulations.

@@

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \citet{Schaerer02}.  The radiation field is evolved with
with adaptive ray tracing \citep{Abel02_RT, Wise10} and is coupled
self-consistently to the hydrodynamics.  These stars die as
pair-instability SNe (PISNe) if they are in the mass range 140--260
\Ms \citep{Heger03}.  We use the explosion energy from
\citet{Heger02}, where we fit the following function to their models,
$E_{\rm PISN} = 10^{51} \times [5.0 + 1.304 (M_{\rm He} - 64)]$, where
$M_{\rm He} = (13/24) \times (M_\star - 20) \Ms$ is the helium core
mass and $M_\star$ is the stellar mass.

@@

Here we present the gaseous and stellar evolution of two selected
halos in the simulation, starting with the birth of the first Pop III
star in its merger history.  We then compare the nature of star
formation in these two halos.

\subsection{Halo evolution}

\li Describe the evolution of the baryon fraction and metallicity from
PISN and Type II SNe metals in the two halos.

@@

\li Report on the Pop III stars forming in halo progenitors

@@

\li Present results on the star formation history and the physical
reasons for the features seen in the SFH and metallicities.

\li Present the stellar metallicity distribution at the final time

\subsection{Environments for starbursts and quiescent star formation}

\li Present the environment for bursting and quiescent star formation
history

@@

Support for this work was provided by NASA through Hubble Fellowship
grant \#120-6370 awarded by the Space Telescope Science Institute,
which is operated by the Association of Universities for Research in
Astronomy, Inc., for NASA, under contract NAS 5-26555.  Computational
resources were provided by NASA/NCCS award SMD-09-1439.  The majority
of the analysis and plots were done with \texttt{yt} \citep{yt}.
